---
title: "Week 04: Regular expressions"
format: 
  revealjs:
     logo: hu.png
     css: logo.css
     slide-number: true
     show-slide-number: all
     preview-links: auto
     theme: default
  pdf: default
editor: visual
footer: "Automated Web Data Collection And Text As Data (Erfort 2023)"
date: "2023-05-09"
date-format: long
from: markdown+emoji
---

```{r install, echo = F}
# install.packages("rmarkdown")
# install.packages("tidyverse")
# install.packages("qrcode")

library(tidyverse)
library(rmarkdown)
library(qrcode)
```

## Plan for today

-   Housekeeping :broom:
-   Research article and discussion
-   Assignment from last week
-   Strings
-   Regular expressions
-   `stringr`
-   Live coding
-   Exercise

## Housekeeping :broom:

-   Any questions organizational questions?
-   Any problems with R or Posit Cloud?
-   Readings

------------------------------------------------------------------------

## Presentation

------------------------------------------------------------------------

## Strings

-   Create a string using quotes

. . .

```{r echo = T}
string1 <- "This is a string"
string2 <- 'If I want to include a "quote" inside a string, I use single quotes'
```

------------------------------------------------------------------------

## Special characters and escapes

-   Special characters have a function and will be interpreted by R in that way
-   The most obvious one is the quote itself "

. . .

-   To tell R that we want it to be interpreted as a normal character, we use the backslash  as an escape

. . .

```{r echo = T}
"This is a quote\""
```

. . .

-   Other characters with special functions: `\*.(){}\[\]-` etc.

------------------------------------------------------------------------

## Regular expressions

-   A way of defining patters in strings
-   Perfect for extracting data (from semi-structured data)
-   You don't need to learn everything! Just a few basics will help us a lot to collect the data we want.

------------------------------------------------------------------------

## Examples for using regular expressions

-   We have a full name in our dataset, but we want the first and last names separately:
    -   `Smith, Charlie` -\> `Smith` + `Charlie`

. . .

-   We want to extract the date from a paragraph of text:
    -   `The article was published on 15 July 2012.` -\> `15 July 2012`

------------------------------------------------------------------------

## The `stringr` package

-   `str_c(character_vector)` combines two or more character vectors

. . .

-   `str_extract(string, pattern)` extracts a part of a string that matches the pattern provided.

. . .

-   `str_detect(string, pattern)` returns TRUE when the pattern is present in the string, otherwise FALSE

. . .

-   `str_remove(), str_extract_all(), str_replace(), str_to_lower(), str_split()`...

------------------------------------------------------------------------

## Basic regex patterns

-   Normal characters only match exactly

. . .

```{r echo = T}
library(babynames)
str_view(fruit, "berry", match = NA)
```

------------------------------------------------------------------------

## Basic regex patterns

-   Normal characters only match exactly

. . .

```{r echo = T}
str_view("automated web data collection", "data", match = NA)
```

------------------------------------------------------------------------

## Punctuation

-   `.`, `+`, `*`, `[`, `]`, and `?` have special meanings.
-   E.g. `.` matches any character

. . .

```{r echo = T}
str_view(c("Humboldt University", "Free University", "Technical University"), ".......University", match = NA)
```

------------------------------------------------------------------------

## Quantifiers

-   Quantifiers control how many times a pattern can match:

-   `?` makes a pattern optional (i.e. it matches 0 or 1 times)

-   `+` lets a pattern repeat (i.e. it matches at least once)

-   `*` lets a pattern be optional or repeat (i.e. it matches any number of times, including 0).

. . .

```{r echo = T}
letters <- c("aaaab", "bbbbccd", "bbaa")

str_extract(letters, "a+")

str_extract(letters, "a?")
```

. . .

```{r echo = T}
str_extract(letters, ".*")
```

## Character classes

-   ... are defined by `[]` and let you match a set of characters
-   `[abc]`, or `[a-z]`, or `[A-Z]`
-   Use `^` to exclude these characters: `[^abc]`

. . .

```{r echo = T}
str_view(c("abcdef123", "abc23", "cdefghik345"), "[abc]", match = NA)
```

. . .

```{r echo = T}
str_view(c("abcdef123", "abc23", "cdefghik345"), "[^0-9]", match = NA)
```

## Anchors

-   Help you match only patterns at the start or end of a string
-   `^` for the start, `$` for the end

. . .

```{r echo = T}
str_view(c("mr charlie smith", "ms charlie palmr"), "mr", match = NA)
```

. . .

```{r echo = T}
str_view(c("mr charlie smith", "ms charlie palmr"), "^mr", match = NA)
```

## More quantifiers

-   We can specify, how often something has to appear using `{}`
    -   `{n}` matches exactly n times.
    -   `{n,}` matches at least n times.
    -   `{n,m}` matches between n and m times.

. . .

```{r echo = T}
str_view(c("aaaa", "aaaaaaa", "aaaaaaaa"), "a{5}", match = NA)
```

## Lookaheads/lookbehinds

-   We can specify a pattern, by saying what pattern comes before/after
-   Positive lookahead: `(=>)`
-   Positive lookbehind: `(?<=)`

. . .

```{r echo = T}
mytexts <- c("I don't want this text." ,"I only want this text. Contact: Twitter @something")
str_view(mytexts, ".*(?=Contact:)", match = NA)
```

. . .

```{r echo = T}
myurls <- c("This is not a URL.", "URL: www.ineedthisurl.com")
str_view(myurls, "(?<=URL: ).*", match = NA)
```

## Live coding

------------------------------------------------------------------------

------------------------------------------------------------------------

## Next week

<br>

::: columns
::: {.column width="50%"}
Readings :book:

-   Automated Data Collection with R
    -   [Ch. 9 (until 9.1.4), and Ch. 9.2.2](https://doi.org/10.1002/9781118834732.ch9)
-   Grolemund and Wickham (2023)
    -   [Ch. 28](https://r4ds.hadley.nz/strings.html)
:::

::: {.column width="50%"}
`Coding` :robot:

-   Exercise 04
-   Play Regexone
-   Munzert et al. 2014, [Ch. 9](https://doi.org/10.1002/9781118834732.ch2), Problems 1-14 (read, don't solve!)
:::
:::

## Tips for the exercises

-   Coding is trial and error!
-   Build your code step by step
-   Google error messages
-   Ask ChatGPT for help
-   Help each other, ask questions on Moodle

------------------------------------------------------------------------

See you next week :wave:
