---
title: "Week 06: APIs, JSON and other file formats"
format: 
  revealjs:
     logo: hu.png
     css: logo.css
     slide-number: true
     show-slide-number: all
     preview-links: auto
     theme: default
  pdf: default
editor: visual
footer: "Automated Web Data Collection And Text As Data (Erfort 2023)"
date: "2023-05-23"
date-format: long
from: markdown+emoji
---

```{r install, echo = F}
# install.packages("rmarkdown")
# install.packages("tidyverse")
# install.packages("qrcode")

library(tidyverse)
library(rmarkdown)
library(qrcode)
library(rvest)
library(xml2)

```

## Plan for today

-   Housekeeping :broom:
-   APIs
-   JSON
-   Other file formats
-   Live coding
-   Exercise

## Housekeeping :broom:

-   Any questions?
-   Assignment from last week

## Recap

First step

```{mermaid}
flowchart LR
  A(URL)-- "read_html(URL)" --> B(HTML file)
  B -- "rvest/regex" --> C(Data)
  B -- "rvest" --> D(list of URLs)
```

or

```{mermaid}
flowchart LR
  A(URL)-- "manually" --> B(URL pattern)
  B -- "str_c(URL, pattern)" --> C(list of URLs)
```

# Recap

Using loops to download multiple HTMLs

```{mermaid}
flowchart LR
  B(URL i)-->C{filename}
  C--"if(file.exists(filename))"-->D("i+1")
  D-.->B
  C--"if not"-->G("read_html() + <br>write_html()")
  G-->H("i+1")
  H-.->B
```

# Recap

Using loops to extract data from multiple HTMLs

```{mermaid}
flowchart LR
  A(multiple files)-- "list.files()"  --> B(list of files)
```


```{mermaid}
flowchart LR
  B(File i)--"rvest/regex"-->C(Data)
  C--"add to"-->D("Dataframe")
  D-->E
  E(i+1)-.->B
```

## APIs

-   Webscraping often uses the user interface (i.e. the visual website)
-   But websites often provide data in a more efficient way
-   Application Programming Interfaces

## APIs

-   Standardized way to request data
-   API returns only the requested data
-   Some APIs are free
-   Some APIs require authentication

## API example

``` {.html code-line-numbers="|1-2|12-15"}
{"latitude":54.100006,
"longitude":12.100006,
"generationtime_ms":0.27310848236083984,
"utc_offset_seconds":0,
"timezone":"GMT",
"timezone_abbreviation":"GMT",
"elevation":13.0,
"daily_units":
  {"time":"iso8601",
  "temperature_2m_max":"°C"},
"daily":
  {"time":
      ["2020-05-16","2020-05-17","2020-05-18","2020-05-19","2020-05-20"],
  "temperature_2m_max":
      [14.0,16.4,13.6,16.0,15.1]}}
```

<https://archive-api.open-meteo.com/v1/archive?latitude=54.09&longitude=12.14&start_date=2020-05-16&end_date=2020-05-17&hourly=temperature_2m>

## APIs more examples

-   Geographic coordinates
-   Twitter data
-   Uber price estimation

## APIs more examples

List of free APIs

```{r }
qr_code("https://github.com/public-apis/public-apis") %>% plot
```

<https://github.com/public-apis/public-apis>

## R packages for APIs

-   Often R packages make it easier to use APIs in R

List of R packages

```{r }
qr_code("https://gist.github.com/zhiiiyang/fc19995f7e350f3c7fb940757f6213cf#file-apis-md") %>% plot
```

<https://gist.github.com/zhiiiyang/fc19995f7e350f3c7fb940757f6213cf#file-apis-md>

## R packages for APIs

-   `rtweet` package to access Twitter API
-   Twitter API requires authentication

## File formats

Which file formats do we need to know?

. . .

-   XML `xml2::read_xml()`
-   CSV `readr::read_csv()`
-   XLSX `openxlsx::read.xlsx()`
-   JSON

. . .

Next week:

-   PDF
-   JPG etc.

## JSON

-   Data are stored in key-value pairs: `{"key": "value"}`
-   JSON is hierarchical, unlike tables/dataframes
-   Curly brackets `{}` define objects
-   Square brackets `[]` define arrays (ordered sequence/list of values)

## JSON

``` {.html code-line-numbers="|1-12|3|5-8|9|13-15"}
{"indy movies" :[
  {
  "name" : "Raiders of the Lost Ark",
  "year" : 1981,
  "actors" : {
      "Indiana Jones": "Harrison Ford",
      "Dr. Rene Belloq": "Paul Freeman"
    },
  "producers": ["Frank Marshall", "George Lucas", "Howard Kazanjian"],
  "budget" : 18000000,
  "academy_award_ve": true
  },
{
"name" : "Another Movie"
}]
}
```

## JSON

```{r echo = T}
library(jsonlite)
json <- '{"indy movies" :[
  {"name" : "Raiders of the Lost Ark",
  "year" : 1981,
  "actors" : {
      "Indiana Jones": "Harrison Ford",
      "Dr. Rene Belloq": "Paul Freeman"}},
{"name" : "Another Movie",
"year" : 1999
}]}'
  
fromJSON(json)
```

## Hierarchical data

-   We usually work with rectangular data.
-   How can we transform hierarchical data to rectangular?

. . .

        
- add additional columns: `tidyr::unnest_wider()`
- add additional rows: `tidyr::unnest_longer()`



## Hierarchical data

-   Packages: `jsonlite` and `tidyr`
-   `jsonlite::parse_json()`
-   `jsonlite::fromJSON()`

## Live coding

------------------------------------------------------------------------

## Takeaways this week

-   R can handle different file formats.
-   When data is not rectangular, we can transform it.

------------------------------------------------------------------------

## Next week

<br>

::: columns
::: {.column width="50%"}
Readings :book:

-   Hoffmann, Mark (2021) -[Chapter 11](https://bookdown.org/markhoff/css/working-with-weird-data-an-example.html)
:::

::: {.column width="50%"}
`Coding` :robot:

-   Exercise 06
:::
:::

## Tips for the exercises

-   Coding is trial and error!
-   Build your code step by step
-   Google error messages
-   Ask ChatGPT for help
-   Help each other, ask questions on Moodle

------------------------------------------------------------------------

See you next week :wave:
